---
title: "Building up an MCMC function"
author: Eric C. Anderson
date: April, 2014
output: 
  html_document:
    toc: true
    highlight: pygments
  pdf_document:
    toc: true
    toc_depth: 3
    highlight: pygments
---

Note that to get this to really be a vignette, I must insert this at the top:
```
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Building up an MCMC function}
-->
```

```{r}
#library(fullsniplings)
```

## Overview of what this function looks like
I am building up the functions for this in the file `mcmc1.R`.  Check that out.

## Some quick benchmarking:
I am doing a little benchmarking.  First we get some data together and some variables prepped:
```{r}
# set values from test data
genos <- fs_dev_test_data$mykiss_genos
mu <- 0.005

# first get all the data. Store it in a list that we will add other Variables to, as well
Vars <- prep_all_variables(genos = genos, geno_error_rates = mu)
# set the full sibling list to be the true full sibships
FSL <- fs_dev_test_data$mykiss_sib_list
# from that list, make the "Individuals' full sibling groups" vector
IFS <- make_IFS_from_FSL(FSL, ncol(Vars$geno_liks))
# set the LMMI to be Vars$pk_marriage_liks (just to keep notation consistent with my notebook notes)
LMMI <- Vars$pk_marriage_liks
# get the initial condition for the LMMFS
LMMFS <- make_LMMFS_from_FSL_and_LMMI(FSL, LMMI)
# get the initial condition for the posteriors on the parent genotypes from each full sibship in LMMFS
PMMFS <- make_PMMFS_from_FSL_and_LMMFS(FSL, LMMFS, Vars$afreqs)
# now we set the initial conditions on the Kid-Prongs!
KidProngs <- make_KidProngs_from_FSL_and_PMMFS(FSL, PMMFS)
```

Now, we are going to compute the full conditional (just the likelihood parts) of individual 13 (base-0 he is 12) 
who is a member of the largest full sibling group (which is first on the FSL).  We do it via R functions and 
also with my Cpp function:
```{r}
DoneByR <- sapply(1:ncol(KidProngs), function(x) prod(colSums(matrix(KidProngs[,x] * Vars$geno_liks[, 13], nrow=3))))
DoneByC <- kid_prongs_times_ind_likelihoods(FSL, Vars$geno_liks[, 13], KidProngs)
cbind(DoneByC, DoneByR[1:length(DoneByC)])[1:10,]
```
Looking at the results, see that clearly the first sibship is the correct one!

Now, let's see how long these take:
```{r}
microbenchmark(kid_prongs_times_ind_likelihoods(FSL, Vars$geno_liks[, 13], KidProngs))
```
That is 675 microseconds for all 803 of the full sibling groups.  So, less than 1/1000th of a second to updatea
against all 803.  That is cool.

### The gibbs update function
Let's try a little test of our Gibbs update function.  It is incomplete, but should spit back those
full conditional likelihoods
```{r}
GFREQS <- gfreqs_from_afreqs(Vars$afreqs)
UGP <- unrelated_pair_gfreqs(gfreqs_from_afreqs(Vars$afreqs))
TP  <- trans_probs()
f <- function(Ind)  gibbs_update_one_indiv_in_place( 
  FSL,
  IFS,        # length N
  LMMI,       # 3 * 3 * L rows, N columns
  LMMFS,      # 3 * 3 * L rows, N columns
  PMMFS,      # 3 * 3 * L rows, N columns
  KidProngs,  # 3 * L rows, N columns
  integer(0),       # IntegerVector treated as a stack
  GFREQS,     # Genotype freqs given the allele freqs 3 rows, L cols
  UGP,        # unrelated_pair_gfreqs(gfreqs_from_afreqs(af))  3 * 3 rows, L cols
  TP,         # Transmission probs 3 * 3 * 3
  Vars$geno_liks,    # Individual geno likes.  3 * L rows, N columns
  Ind                   # Index (base-0) of individual to be updated
)
```
Now we have defined ``f(x)`` to do what turns out to be most of the computations for an update to individual x.
```{r}
microbenchmark(f(12))
```