
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{data for testing during development}
-->


Background on data sets for tests during development
========================================================

This is just a little background on how and why I went about creating a few
small data sets and variables and such that I have used in the course of 
testing while developing this package.

It is not intended that the test data be actually generated by knitting this
document while building.  But it certainly could be run, and an assignment made
and saved to the `data` directory.

The data that is described here has all be saved into the list `fs_dev_test_data` which 
we will declare here:
```{r}
fs_dev_test_data <- list()
```

## Simple SNP data set
Here we want something that we can use to test that our data input is working:
```{r, cache=TRUE}
# make SNPs that have two of various symbols that might be used
# to denote alleles
set.seed(5)
num.inds <- 20
dat <- lapply(list(
  c("A","G"), 
  c("T", "C"),
  c(1,2),
  c("X", "Y"),
  c("Vic", "Fam")),
    function(x) cbind(sample(x, size=num.inds, replace=T),
                      sample(x, size=num.inds, replace=T) )   
)
# add some missing data:
dat <- lapply(dat, function(x) {x[sample(1:num.inds, size=3, replace=F), ] <- NA; x})
dat <- as.data.frame(dat, stringsAsFactors=F)
colnames(dat) <- make.unique(paste("Locus", rep(1:5, each=2), sep=""))
rownames(dat) <- paste("Ind", 1:num.inds, sep="")
fs_dev_test_data$plain_snp_data <-  dat
```


## Big Mykiss SNP-data set
I want to add in the SNP data that I used when Hans and I were looking at the the `psiblings` method.  It is 
a nice size, about 1400 fish at 95 SNPs or so.  We will save the genotype data and the pedigree.  This is a little
ugly now, and not portable, because the original is on my hard drive.  I must clean this up before distribution,
but I wanted a record of it for now.
```{r}
# get the data
load("/Users/eriq/Documents/work/prj/HansSib/PBT_Data/WSH/data/wsh_data_for_sibship_analysis.Rda")
fs_dev_test_data$mykiss_pedigree <- ped.no.parents
fs_dev_test_data$mykiss_genos <- genos.no.parents
names(fs_dev_test_data$mykiss_genos) <- make.unique(paste("Loc_", rep(1:(ncol(genos.no.parents)/2), each=2), sep=""))
```
}

### Making a full-sib list from the mykiss pedigree
First we will disregard half siblings and just make a list of of integer identifiers of the full siblings.  
```{r}
# here are the indices of the kids starting from 0
kidx <- as.integer(factor(fs_dev_test_data$mykiss_pedigree$Kid, levels=as.character(fs_dev_test_data$mykiss_pedigree$Kid))) - 1
sib_list <- split(kidx, paste(fs_dev_test_data$mykiss_pedigree$Pa, fs_dev_test_data$mykiss_pedigree$Ma, sep="--"))
ord_sib_list <- sib_list[order(sapply(sib_list, length), -sapply(sib_list, function(x) x[1]), decreasing=T)]
fs_dev_test_data$mykiss_sib_list <- ord_sib_list
```


## Big Chinook SNP-data set
I am going to add one of the chinook data sets in which we inferred siblings from PBT.  I really should get a more updated one from
Anthony, but for now I am just going to use what I had from the HansSib project.
```{r}
load("/Users/eriq/Documents/work/prj/HansSib/PBT_Data/FRH/FRH_sib-data.Rda")
chinook_pedigree <- as.data.frame(matrix(unlist(strsplit(rownames(genos), split = "--")), byrow=T, ncol=3), stringsAsFactors = F)
names(chinook_pedigree) <- c("Kid", "Pa", "Ma")

# put the names in there as just the kids
chinook_genos <- genos
rownames(chinook_genos) <- chinook_pedigree$Kid

# here are the number of full sibships each parent is in:
num.fss <- table(unlist(strsplit(unique(paste(chinook_pedigree$Ma, chinook_pedigree$Pa, sep="--")), "--")))

# here are the parents that are in more than one full sib group:
hssers <- names(num.fss)[num.fss>1] 

# these are the indices of the kids to keep as they are all full sibs,
# (no half sibling relationships in there that we no of)
retain <- !((chinook_pedigree$Ma %in% hssers) | (chinook_pedigree$Pa %in% hssers) )

# now, make data sets that should have no half sibs in them
chinook_full_sibs_pedigree <- chinook_pedigree[retain, ]
chinook_full_sibs_genos <- chinook_genos[retain, ]

fs_dev_test_data$chinook_genos <- chinook_genos
fs_dev_test_data$chinook_pedigree <- chinook_pedigree
fs_dev_test_data$chinook_full_sibs_genos <- chinook_full_sibs_genos
fs_dev_test_data$chinook_full_sibs_pedigree <- chinook_full_sibs_pedigree
```

## Save it all into a data object at the end
Though I actually have this commented out:
```{r}
# save(fs_dev_test_data, file="data/fs_dev_test_data.rda", compress="xz")
```