
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{data for testing during development}
-->


Background on data sets for tests during development
========================================================

This is just a little background on how and why I went about creating a few
small data sets and variables and such that I have used in the course of 
testing while developing this package.

It is not intended that the test data be actually generated by knitting this
document while building.  But it certainly could be run, and an assignment made
and saved to the `data` directory.

The data that is described here has all be saved into the list `fs_dev_test_data` which 
we will declare here:
```{r}
fs_dev_test_data <- list()
```

## Simple SNP data set
Here we want something that we can use to test that our data input is working:
```{r, cache=TRUE}
# make SNPs that have two of various symbols that might be used
# to denote alleles
set.seed(5)
num.inds <- 20
dat <- lapply(list(
  c("A","G"), 
  c("T", "C"),
  c(1,2),
  c("X", "Y"),
  c("Vic", "Fam")),
    function(x) cbind(sample(x, size=num.inds, replace=T),
                      sample(x, size=num.inds, replace=T) )   
)
# add some missing data:
dat <- lapply(dat, function(x) {x[sample(1:num.inds, size=3, replace=F), ] <- NA; x})
dat <- as.data.frame(dat, stringsAsFactors=F)
colnames(dat) <- make.unique(paste("Locus", rep(1:5, each=2), sep=""))
rownames(dat) <- paste("Ind", 1:num.inds, sep="")
fs_dev_test_data$plain_snp_data <-  dat
```


## Save it all into a data object at the end
Though I actually have this commented out:
```{r}
# save(fs_dev_test_data, file="data/fs_dev_test_data.rda")
```