---
title: "Full-Sibling Chinook MCMC"
author: Eric C. Anderson
date: April, 2014
output: 
  html_document:
    toc: true
    highlight: pygments
  pdf_document:
    toc: true
    toc_depth: 3
    highlight: pygments
---

Note that to get this to really be a vignette, I must insert this at the top:
```
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Running on chinook data and comparing to Colony}
-->
```

```{r}
library(fullsniplings)
```

## Intro
So, are going to use the chinook full sibling data:
```{r}
dim(fs_dev_test_data$chinook_full_sibs_genos)
```
That is `r nrow(fs_dev_test_data$chinook_full_sibs_genos)` at `r ncol(fs_dev_test_data$chinook_full_sibs_genos)/2` SNPs.

## Running the chain
We can run the chain like this:
```{r}
#chinook_results <- run_mcmc(fs_dev_test_data$chinook_full_sibs_genos, burn_in = 50, num_sweeps = 200)
#save(chinook_results, file="chinook_chain_results.rda")

# i have commented that out so it doesn't take took much time, and I can get it back quickly
# like this:
load("/Users/eriq/Documents/git-repos/fullsniplings/chinook_chain_results.rda") # not portable file name
# but render() in rmarkdown seems to make the directory change...
```

## Getting the truth
Here I pull out some code from `observe_mcmc.R` to get the hashable names of the
different true sibling groups as a character vector called `true_sibgroups_as_strings`.
```{r}
ped <- fs_dev_test_data$chinook_full_sibs_pedigree
ordered.kids <- rownames(fs_dev_test_data$chinook_full_sibs_genos)
ped$kididx <- as.integer(factor(ped$Kid, levels = ordered.kids)) - 1
kid.split <- split(ped$kididx, paste(ped$Pa, "---", ped$Ma, sep=""))

# here we get a list of all the sibling groups
sibgroups <- unname(kid.split[order(sapply(kid.split, length), decreasing=T)])

# and here we can store them as strings:
true_sibgroups_as_strings <- sapply(sibgroups, function(x) paste(x, collapse="-"))
```

And, I also get a list of all the siblings that belong to each individual, in the truth:
```{r}
# now get a list of the siblings of an individual
sg.list <- vector("list", length = nrow(ped))
for(L in kid.split) {
  for(ind in L) {
    sg.list[[ind+1]] <- setdiff(L, ind)
  }
}
```
Now, if `Ind` is the base-0 index of an individual then
`sg.list[[Ind+1]]` is a vector of the base-0 indices of all of its full siblings


## Assessing results compared to the truth
Put the greedy partition results into a table with a short name, then add some columns to it:
```{r}
crtab <- chinook_results$Partition
crtab$Correct <- rownames(crtab) %in% true_sibgroups_as_strings
crtab$CumulCorrectSibs <- cumsum(crtab$NumSibs * crtab$Correct)
crtab$CumulSibsInIncorrectSibships <- cumsum(crtab$NumSibs * (1-crtab$Correct))
```
Note that the last line gives us a really punishing measure: if a sibship is incorrect, even if it is only
off by one individual, everyone in it is counted as wrong.  

There is one sibship with posterior of 1 and it is size 8, but it is wrong.  Compare these:
```{r}
# here is the row that has them
crtab["94-105-231-405-529-532-551-847", ]

# and here is what the sibship probably should look like:
c(94, sg.list[[95]])
```
I think we have seen these guys before and they are full cousins or something.

### Testing whether there are any non-sibs, etc.
We will add some more columns to the table
```{r}
# this is a list of sibgroups by their base-0 index
S <- lapply(strsplit(rownames(chinook_results$Partition), "-"), as.numeric)
OmAndCom <- ave_err_commission_and_ommission(S, sg.list)
crtab$AveNumNonSibs <- OmAndCom$AveNumNonSibs
crtab$AveNumSibsMissing <- OmAndCom$AveNumSibsMissing
```



## Getting Colony's results
This will be a little weird.  I will use a previous Colony run, since I don't have time to re-run it.  That
previous run included 50 extra individuals (some half-siblings) which I will discard from the results which 
will give us something that is almost comparable.

### Running Time
First I note that I can look at the standard output:
```
2014-05-29 11:53 /PairX1/--%  pwd
/Users/eriq/Documents/xp_dev_svn_checkouts/slg_pipe/arena/FRH_ColonyRun/ColonyArea/Collections/AA/PairX1
2014-05-29 11:55 /PairX1/--% gzcat StdoutColony.txt.gz 

# and find in there the running time:
TotalRunningTime(Min)=1.458552E+02
```
So, running time is 146 minutes, which I can call about 2.5 hours, and that, remember is for the *pairwise*
version of Colony.


### Colony-inferred sibships

#### First get the colony inferred sibships and put their long names on them
```{r}
# read in the colony sibships
colony_sib_file <- system.file("data_files/colony/AA_PairX1.BestFSFamily", package="fullsniplings", mustWork = T)
csibs <- SlurpSibships(file=colony_sib_file, skip=1)

# now, put the names back on them:
load(system.file("data_files/colony/FRH-sibs-for-pipeline.Rda", package="fullsniplings", mustWork = T))
names(fish.names) <- rownames(data)
idxs <- 1:nrow(data)
names(idxs) <- rownames(data)
csibs <- NamizeSibs(csibs, fish.names, idxs)
```


#### Filter it down to just the full sibs I retained for fullsniplings' analysis
At this juncture the names of these fish in the colony output are like this: `T065899_F_10-6-2010--T030068_?_9-28-2007--T030092_?_9-28-2007`.
So, we can use what is in `ped` to go immediately to the base-0 index of each fish, just like we have for the 
full-sniplings output:
```{r}
bz_idx <- ped$kididx
# make names for this base-0 index vector by catenating Kid-Pa-Ma
names(bz_idx) <- paste(ped$Kid, ped$Pa, ped$Ma, sep="--")

# and then make a sibship list of just those indices:
colony_sibs <- lapply(csibs$sibs.names, function(x) {z <- unname(bz_idx[x]); z[!is.na(z)]})
# drop the empty ones
empties <- sapply(colony_sibs, length)>0
colony_sibs <- colony_sibs[empties] 
colony_probs <- as.numeric(csibs$probs[empties])
```

And now we want to make a data frame of those results that is similar to `crtab`
```{r}
colony_tab <- data.frame(Posterior = colony_probs, NumSibs = sapply(colony_sibs, length))
rownames(colony_tab) <- sapply(colony_sibs, function(x) paste(sort(x), collapse="-"))
# now put them in order by posterior and then size
ord <- order(colony_tab$Posterior, colony_tab$NumSibs, decreasing = T)
colony_tab <- colony_tab[ord, ]

colony_tab$Correct <- rownames(colony_tab) %in% true_sibgroups_as_strings
colony_tab$CumulCorrectSibs <- cumsum(colony_tab$NumSibs * colony_tab$Correct)
colony_tab$CumulSibsInIncorrectSibships <- cumsum(colony_tab$NumSibs * (1-colony_tab$Correct))

# then get average errors of commission and omission
OmAndCom <- ave_err_commission_and_ommission(colony_sibs, sg.list)
colony_tab$AveNumNonSibs <- OmAndCom$AveNumNonSibs
colony_tab$AveNumSibsMissing <- OmAndCom$AveNumSibsMissing
```


## Make some plots
Here is one for my method
```{r}
#quartz(width=10, height=6)
sib_result_plot(crtab$NumSibs, crtab$Visits/200, false_positive = crtab$AveNumNonSibs>0.001, lscale = .03, correct = crtab$Correct)
dev.copy2pdf(file="fullsniplings-posterior-plot.pdf")
```
and here is one for colony:
```{r}
sib_result_plot(colony_tab$NumSibs, colony_tab$Posterior, false_positive = colony_tab$AveNumNonSibs>0.001, lscale = .03, correct = colony_tab$Correct)
dev.copy2pdf(file="colony-posterior-plot.pdf")
#sib_result_plot(crtab$NumSibs, (crtab$Visits/200) - .4, false_positive = crtab$AveNumNonSibs>0.001, lscale = .03, correct = crtab$Correct, add=T)
```


How about some plots of cumulative number of individuals put in the right sibship
and number put in the wrong sibship:
```{r}
plot(colony_tab$CumulCorrectSibs, colony_tab$CumulSibsInIncorrectSibships, 
     xlim = c(0, max(crtab$CumulCorrectSibs)), type = "l", lwd = 2,
     ylab = "Number of Fish Placed in Incorrect or Incomplete Sibling Groups",
     xlab = "Number of Fish Placed in Correct Sibing Groups")
lines(crtab$CumulCorrectSibs, crtab$CumulSibsInIncorrectSibships, lwd = "2", col = "orange")
legend("topleft", 
         legend = c("COLONY", "FULLSNIPLINGS"), 
         col = c("black", "orange"), lty = "solid", lwd = 2
         )
text(x = 0, y = 175, "Totals:" )
text(x = 0, y = 150, paste("FULLSNIPLINGS:    ", crtab[nrow(crtab), "CumulCorrectSibs"], "correct   and   ", crtab[nrow(crtab), "CumulSibsInIncorrectSibships"], "incorrect."), pos = 4  )
text(x = 0, y = 125, paste("COLONY:       ", colony_tab[nrow(colony_tab), "CumulCorrectSibs"], "correct   and   ", colony_tab[nrow(colony_tab), "CumulSibsInIncorrectSibships"], "incorrect."), pos = 4  )
dev.copy2pdf(file = "cumul-correct-vs-incorrect.pdf")
```

We can also think about the cumulative number if individuals that have been placed
in sibships that have at least one non sib, as a function of the total number of 
individuals placed correctly into their sibships.
```{r}
crcum <- cumsum(crtab$NumSibs * (crtab$AveNumNonSibs > 0.01))
colcum <- cumsum(colony_tab$NumSibs * (colony_tab$AveNumNonSibs > 0.01))
plot(crtab$CumulCorrectSibs, crcum,
     type = "l", lwd=2, col = "orange",
     ylab = "Number of Fish Grouped With Non-Siblings",
     xlab = "Number of Fish Placed in Correct Sibing Groups")
lines(colony_tab$CumulCorrectSibs, colcum, lwd = 2)

legend("topleft", 
         legend = c("COLONY", "FULLSNIPLINGS"), 
         col = c("black", "orange"), lty = "solid", lwd = 2
         )

text(x = 0, y = 48, "   Totals:" )
text(x = 0, y = 44, paste("FULLSNIPLINGS:    ", crtab[nrow(crtab), "CumulCorrectSibs"], "correct   and   ", max(crcum), "grouped with a non-sib"), pos = 4  )
text(x = 0, y = 38, paste("COLONY:       ", colony_tab[nrow(colony_tab), "CumulCorrectSibs"], "correct   and   ", max(colcum), "grouped with a non-sib"), pos = 4  )
dev.copy2pdf(file = "cumul-correct-vs-grouped-with-non-sib.pdf")
```

And also cumulative number of individuals placed in sibgroups that are missing at 
least one of their true sibs:
```{r}
plot(crtab$CumulCorrectSibs, cumsum(crtab$NumSibs * (crtab$AveNumSibsMissing > 0.01)),
     type = "l", lwd=2, col = "orange", ylim=c(0,sum(colony_tab$NumSibs * (colony_tab$AveNumSibsMissing > 0.01))))
lines(colony_tab$CumulCorrectSibs, cumsum(colony_tab$NumSibs * (colony_tab$AveNumSibsMissing > 0.01)))
```