---
title: "Quick overview of use/flow of functions"
author: Eric C. Anderson
date: April, 2014
output: 
  html_document:
    toc: true
    highlight: pygments
  pdf_document:
    toc: true
    highlight: pygments
---

Note that to get this to really be a vignette, I must insert this at the top:
```
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{quick overview of use/flow of functions}
-->
```

I am just writing this down while I am writing these functions, to remind me of how I am going to use them.

## Alleles and allele frequencies
Here are the standard steps that I would take in prepping up all the variables.  You will of
course want to prepare by loading the library:
```{r}
library(fullsniplings)
```
### First, turn a data frame of SNP alleles into values of 0, 1, 2, or NA for each individual: 
```{r}
snp_genos <- get_snp_genos(fs_dev_test_data$plain_snp_data)
snp_genos
```

### Second, turn those 0, 1, 2, or NA genotypes into indicator vectors.  
So a 0 would be (1,0,0), a 1 would be (0,1,0), a 2 would be (0,0,1), and an NA would be (NA, NA, NA)
```{r}
snp_indics <- genos_to_indicators(g = snp_genos$mat)
snp_indics[, , 1:3]
```
Note that this creates a 3-D array.

### Third, count up the number of observed genotypes:
```{r}
geno_counts <- count_genos(snp_indics)
geno_counts
```

### Fourth, compute the allele freqs
```{r}
afreqs <- alle_freqs(geno_counts)
afreqs
```

## Genotype likelihoods for each individual and "marriages"
Getting these likelihoods is pretty straightforward.  They **don't** depend on the allele frequencies.  These are just the
probability of an individuals observed genotype given the (typically unknown) true genotype.  This, of course, depends on the
genotyping error rate of each locus.

### First, compute probs under genotyping error
We make a big 3 x 3 x L array of the probs of observing 0, 1, or 2, (the columns) given the true genotype is 0, 1, or 2 (in the rows).  Under our simple independent mutation model that we are using, where each gene copy has a chance $\mu$ of being detected incorrectly, our likelihood table for $P(\text{Observed geno}=y~|~\text{True geno}=x)$ for a single locus looks like this:
\[
\begin{array}{lr|ccc}
& & & y & \\\\
&                             &      0           &          1           &     2         \\\\ 
\hline
&            0                &   (1-\mu)^2      &     2\mu(1-\mu)      &   \mu^2       \\\\
x~~&        1                &   \mu(1-\mu)     &    \mu^2 + (1-\mu)^2 &   \mu(1-\mu)  \\\\
&            2                &   \mu^2          &     2\mu(1-\mu)      &   (1-\mu)^2  \\\\
\end{array}
\]
It is important to stress that this is just a super simple genotyping error model.  However, 
*any* error model that you might dream up (so long as it is independent between loci) 
can be parameterized by a matrix like this.  So, if you want a different genotyping error
model, it can be specified like this.  Note that we can probably estimate the elements of
this matrix with very large samples.

We compute this in the function `lik_array_from_simple_geno_err()`, and though we think of the
result as a 3 x 3 x L array, we actually return it as a 3 x 3*L matrix as that will facilitate
matters later.

Here is an example of what it looks like to run it for three loci at three different
genotyping error rates:
```{r}
lik_array_from_simple_geno_err(L = 3, mu = c(0.0, .005, .1))
```
In the output, notice that the column names are such that `mu_1.0` describes the situation in
which one observes a genotype of 0 at a locus with genotyping error rate  `mu_1`.


### Next, get the genotype likelihoods for all the individuals
To actually get genotype likelihoods for each of the individuals, given their
observed genotypes, we call the function `get_indiv_geno_lik()` which calls the function 
`lik_array_from_simple_geno_err`.  There is some funky stuff to deal with NAs here, and 
a little bit of weird vectorization, but it seems to be working nicely.  As long as it doesn't
gobble up too much memory with really large data sets.

Here it is in action:
```{r}
geno_liks <- get_indiv_geno_lik(
                      SI = snp_indics, 
                      mu = c(0.0, 0.001, 0.005, 0.01, 0.1)
                      )
```
Let's look at the result for the first four individuals:
```{r}
geno_liks[,,1:4]
```
And compare that to their observed genotypes to see how it all works together.
```{r}
snp_genos$mat[,1:4]
```
Notice in particular, that the likelihood for the true genotype of an individual at a locus
for which data is missing at that individual is constant (1).

### Next, we also compute and store the likelihoods of each individual's pair of parents (i.e. for the "marriage" that produced the individual)
In the immediately preceding section we did a calculation that yielded the likelihood that an individual's
true genotype is 0, 1, or 2, given the observed genotype of that individual.  Here are ambitions are one generation greater.
We want to compute the likelihood of the genotypes of the two parents of each individual given the individual's observed
genotype.  There are 9 possible genotypes of the pair (Paren One can be 0, 1, or 2, and so can parent Two.).  These could be 
condensed to 6 equivalence classes, but that is not worth it here, I feel.  So we won't.  Here are some example probabilities
that we are dealing with:

Parent One  |  Parent Two   |   Kid       |  Prob
----------- | ------------  |  --------   |  ------
0           | 0             | 0           |  1
0           | 0             | 1           |  0
0           | 0             | 2           |  0
0           | 1             | 0           |  $\frac{1}{2}$
0           | 1             | 1           |  $\frac{1}{2}$
1           | 1             | 0           |  $\frac{1}{4}$

and so forth.

#### Create an array that holds those probabilities
We only need to make a single 3x3x3 array for these.  Rows are Parent One, columns are Parent Two,
and the next dimension is for Kids.
We just have a simple function that fills such an array. Check out its output:
```{r}
trans_probs()
```

#### Then use the above probabilities to compute the likelihoods of the marriages that produced each offspring, given the offspring observed genotype
This is something that can probably be done efficiently in R with some wrangling, but guessing that it will be easier in compiled
code with some for loops, I did it with Rcpp.

```{r}
tprobs <- trans_probs()
```